---
title: "Lab 03 - Nobel laureates"
author: "Allison Li"
date: "01/29/25"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

### Exercise 1.1

```{r variables}
as_tibble(nobel)
print(nobel)
```

There are `r nrow(nobel)` observations and `r ncol(nobel)` variables in the dataset. Each row in the nobel dataset represents a Nobel Prize winner. That row includes all the information about the person, such as name, year of winning, and the category of the prize in each column. 

### Exercise 1.2 Create a new data frame called nobel_living

```{r}
nobel_living <- filter(nobel,  is.na(died_date), !is.na(country), gender != "org")
nobel_living
```

There are `r nrow(filtered_data)` observations. 

### Exercise 2

```{r}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )

nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
print(nobel_living_science)

```

### Exercise 3

```{r}

library(ggplot2)

ggplot(nobel_living_science, aes(x = country_us, fill = category)) +
  geom_bar() + 
  coord_flip() +
  facet_wrap(~ category) +
  labs(title = "Nobel Prizes by Category and Country",
       x = "Country", y = "Number of Prizes") 
```


### Exercise 5

...

### Exercise 6

...
