---
title: "Lab 03 - Nobel laureates"
author: "Allison Li"
date: "01/29/25"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

### Exercise 1

```{r variables}
as_tibble(nobel)
print(nobel)
```

There are `r nrow(nobel)` observations and `r ncol(nobel)` variables in the dataset. Each row in the nobel dataset represents a Nobel Prize winner. That row includes all the information about the person, such as name, year of winning, and the category of the prize in each column. 

### Exercise 2 Create a new data frame called nobel_living that filters fo

```{r}
filtered_data <- filter(nobel,  is.na(died_date), !is.na(country), gender != "org")
filtered_data
```

there are `r nrow(filtered_data)` observations. 

### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
