---
title: "Lab 03 - Nobel laureates"
author: "Allison Li"
date: "01/29/25"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

### Exercise 1.1

```{r variables}
as_tibble(nobel)
print(nobel)
```

There are `r nrow(nobel)` observations and `r ncol(nobel)` variables in the dataset. Each row in the nobel dataset represents a Nobel Prize winner. That row includes all the information about the person, such as name, year of winning, and the category of the prize in each column. 

### Exercise 1.2 Create a new data frame called nobel_living

```{r}
nobel_living <- filter(nobel,  is.na(died_date), !is.na(country), gender != "org")
nobel_living
```

There are `r nrow(filtered_data)` observations. 

### Exercise 2

```{r novbel living science}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )

nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
print(nobel_living_science)

```

### Exercise 3

```{r Country}

library(ggplot2)

ggplot(nobel_living_science, aes(x = country_us, fill = country_us)) +
  geom_bar() + 
  coord_flip() +
  facet_wrap(~ category) +
  labs(title = "Nobel Prizes by Category and Country",
       x = "Country", y = "Number of Prizes") 
```

According to the plot graphs, the data supported the Buzzfeed's claim that most living Nobel laureates were based in the US when they won their prizes. For all four types of prizes, most of the living nobel laureates won their prizes when they are in USA. 

### Exercise 4

```{r}
nobel_living_science <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other") 
  )

nobel_living_science %>%
  count(country_us)
```

In total, 147 of the winners are born in the USA. 
### Exercise 5

```{r second variable added}

nobel_living_science <- nobel_living %>%
  mutate(
    born_us = if_else(born_country == "USA", "USA", "Other") 
  )%>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other") 
  )

ggplot(nobel_living_science, aes(x = country_us, fill = born_us)) +
  geom_bar() +  
  coord_flip() +
  facet_wrap(~ category) +  
  labs(title = "Nobel Prizes by Category, Award Country, and Birth country",
       x = "Country", 
       y = "Number of Prizes",
       fill = "Born in US"
  )
```

According to the graphs, although most of the living nobel laureates are based in the US when they won the prizes, most of them in general were not born in the USA. 

### Exercise 6